---
title: "MaStat_thesis_KDN_code"
author: "Kobe De Neve"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, results = FALSE, message = FALSE, warning = FALSE}
# load all required packages
library(tidyverse)
library(ggbiplot)
library(e1071)
library(gridExtra)
library(GGally)
library(flextable)
library(mgcv)
library(performance)
library(mgcViz)
library(ggeffects)
library(visreg)
```

# 1. Loading and Inspecting the Data

```{r, results = FALSE, message = FALSE, warning = FALSE}
# load main data and filter our French source texts
dat_main <- read.delim("data/DATA4.txt",
                       stringsAsFactors = TRUE)

dat_main <- subset(dat_main, lang4 != "FRA")
dat_main <- droplevels(dat_main)

str(dat_main)
View(dat_main)
```

```{r, results = FALSE, message = FALSE, warning = FALSE}
# load dataset with text-level readability measures
dat_readibl <- read.delim("data/READIBL3.txt",
                          stringsAsFactors = TRUE)

str(dat_readibl)
View(dat_readibl)
```

```{r, results = FALSE, message = FALSE, warning = FALSE}
# combine the datasets
dat <- merge(dat_main, dat_readibl,
             by.x = "fileID", by.y = "filename",
             all = TRUE)

dat <- subset(dat, nr_TOT >= 3)
dat <- droplevels(dat)

str(dat)
View(dat)
```

# 2. Exploratory Analysis

## 2.1. Principal Component Analysis

```{r, results = FALSE, message = FALSE, warning = FALSE}
# dat_readibl with lang4 column added
dat_pca <- dat %>%
  select(c(1, 11, 21:81)) %>%
  distinct

# perform pca
pca <- prcomp(dat_pca[, -c(1:2)], center = TRUE, scale = TRUE)
summary(pca)
```

```{r}
# screeplot
ppi = 200
png("plots/pca_screeplot.png", 
    width = 7*ppi, height = 4*ppi, res = ppi)
ggscreeplot(pca, size = 2, linewidth = 0.5) +
  theme_minimal() # elbow at PC4
dev.off()
```

```{r}
# plot of loadings PC1
pca_loadings <- data.frame(pca$rotation)
pca_loadings <- within(pca_loadings, vars <- row.names(pca_loadings))

ppi = 200
png("plots/pc1_loadings.png", 
    width = 7*ppi, height = 8*ppi, res = ppi)
ggplot(pca_loadings, aes(y = vars, x = PC1)) +
  geom_bar(stat = "identity") +
  labs(x = "Contribution to PC1", y = "Variable") +
  theme_minimal()
dev.off()
```

```{r}
# plot of loadings PC2
ppi = 200
png("plots/pc2_loadings.png", 
    width = 7*ppi, height = 8*ppi, res = ppi)
ggplot(pca_loadings, aes(y = vars, x = PC2)) +
  geom_bar(stat = "identity") +
  labs(x = "Contribution to PC2", y = "Variable") +
  theme_minimal()
dev.off()
```

```{r}
# plot of loadings PC3
ppi = 200
png("plots/pc3_loadings.png", 
    width = 7*ppi, height = 8*ppi, res = ppi)
ggplot(pca_loadings, aes(y = vars, x = PC3)) +
  geom_bar(stat = "identity") +
  labs(x = "Contribution to PC3", y = "Variable") +
  theme_minimal()
dev.off()
```

```{r}
# plot of loadings PC4
ppi = 200
png("plots/pc4_loadings.png", 
    width = 7*ppi, height = 8*ppi, res = ppi)
ggplot(pca_loadings, aes(y = vars, x = PC4)) +
  geom_bar(stat = "identity") +
  labs(x = "Contribution to PC4", y = "Variable") +
  theme_minimal()
dev.off()
```

```{r}
# biplot
ppi = 200
png("plots/pca_biplot.png", 
    width = 7*ppi, height = 7*ppi, res = ppi)
ggbiplot(pca, var.axes = FALSE, groups = dat_pca$lang) +
  labs(x = "Standardised PC1 (29.8%)",
       y = "Standardised PC2 (17.6%)",
       color = "Variety") +
  scale_color_discrete(labels = c("Interpreted Dutch",
                                  "Netherlandic Dutch",
                                  "Belgian Dutch")) +
  theme_minimal()
dev.off()
```

## 2.2. Univariate Descriptives

**number of N-grams**

```{r}
# histogram
dat <- within(dat, nr_Ngr <- nr_3gr + nr_4gr)

nr_Ngr_hist <- ggplot(dat, aes(x = nr_Ngr)) +
  geom_histogram() +
  geom_vline(aes(xintercept = mean(nr_Ngr),
                 color = "mean")) +
  geom_vline(aes(xintercept = median(nr_Ngr),
                 color = "median")) + 
  scale_color_manual(
    name = "Statistic",
    values = c("mean" = "blue", "median" = "red")
  ) +
  labs(x = "Number of N-grams") +
  theme_minimal()

# boxplot
nr_Ngr_box <- ggplot(dat, aes(y = nr_Ngr)) +
  geom_boxplot() +
  labs(y = "Number of N-grams") + 
  theme_minimal()

ppi = 200
png("plots/nr_Ngr_uni_des.png", 
    width = 8*ppi,height = 4*ppi, res = ppi)
grid.arrange(nr_Ngr_hist, nr_Ngr_box, nrow = 1,
             top = "Distribution of number of N-grams")
dev.off()
```

```{r}
# summary stats
dat %>%
  summarise(
    mean = mean(rt_FRM),
    median = median(rt_FRM),
    sd = sd(rt_FRM),
    IQR = IQR(rt_FRM),
    skew = skewness(rt_FRM),
    kurtosis = kurtosis(rt_FRM)
  )
```

**rate of *uh(m)***

```{r}
# variable for the rate of 'uh(m)'
dat <- within(dat, rt_EUS <- nr_EUS / nr_TOT)

# histogram
rt_EUS_hist <- ggplot(dat, aes(x = rt_EUS)) +
  geom_histogram() +
  geom_vline(aes(xintercept = mean(rt_EUS),
                 color = "mean")) +
  geom_vline(aes(xintercept = median(rt_EUS),
             color = "median")) +
  scale_color_manual(
    name = "Statistic",
    values = c("mean" = "blue", "median" = "red")
  ) +
  labs(x = "Rate of 'uh(m)'") +
  theme_minimal()

# boxplot
rt_EUS_box <- ggplot(dat, aes(y = rt_EUS)) +
  geom_boxplot() +
  labs(y = "Rate of 'uh(m)'") + 
  theme_minimal()

ppi = 200
png("plots/rt_EUS_uni_des.png", 
    width = 8*ppi, height = 4*ppi, res = ppi)
grid.arrange(rt_EUS_hist, rt_EUS_box, nrow = 1,
             top = "Distribution of disfluency (rate of 'uhm')")
dev.off()
```

```{r}
# summary stats
dat %>%
  summarise(
    mean = mean(rt_EUS),
    median = median(rt_EUS),
    sd = sd(rt_EUS),
    IQR = IQR(rt_EUS),
    skew = skewness(rt_EUS),
    kurtosis = kurtosis(rt_EUS)
  )
```

**language variety**

```{r}
# barplot at utterance level
lang4_utterance <- ggplot(dat, aes(x = lang4)) +
  geom_bar() +
  labs(x = "Variety of Dutch") +
  theme_minimal()

# barplot at text level
lang4_text <- ggplot(dat_pca, aes(x = lang4)) +
  geom_bar() +
  labs(x = "Variety of Dutch") +
  theme_minimal()

ppi = 200
png("plots/lang4_uni_des.png", 
    width = 6*ppi, height = 5*ppi, res = ppi)
grid.arrange(lang4_utterance, lang4_text, nrow = 2,
             top = "Utterance-level (top) and text-level (bottom) distribution of language variety")
dev.off()
```

```{r}
# utterance-level summary tables
with(dat, table(lang4))
with(dat, prop.table(table(lang4)))

# text-level summary tables
with(dat_pca, table(lang4))
with(dat_pca, prop.table(table(lang4)))
```

**lexical density**

```{r}
# histogram
rt_LEX_hist <- ggplot(dat, aes(x = rt_LEX)) +
  geom_histogram() +
  geom_vline(aes(xintercept = mean(rt_LEX),
                 color = "mean")) +
  geom_vline(aes(xintercept = median(rt_LEX),
                 color = "median")) +
  scale_color_manual(
    name = "Statistic",
    values = c("mean" = "blue", "median" = "red")
  ) +
  labs(x = "Lexical density") +
  theme_minimal()

# boxplot
rt_LEX_box <- ggplot(dat, aes(y = rt_LEX)) +
  geom_boxplot() +
  labs(y = "Lexical density") + 
  theme_minimal()

ppi = 200
png("plots/rt_LEX_uni_des.png", 
    width = 8*ppi, height = 4*ppi, res = ppi)
grid.arrange(rt_LEX_hist, rt_LEX_box, nrow = 1,
             top = "Distribution of lexical density")
dev.off()
```

```{r}
# summary stats
dat %>%
  summarise(
    mean = mean(rt_LEX),
    median = median(rt_LEX),
    sd = sd(rt_LEX),
    IQR = IQR(rt_LEX),
    skew = skewness(rt_LEX),
    kurtosis = kurtosis(rt_LEX)
  )
```

**syntactic complexity**

```{r}
# histogram
rt_HYP_hist <- ggplot(dat, aes(x = rt_HYP)) +
  geom_histogram() +
  geom_vline(aes(xintercept = mean(rt_HYP),
                 color = "mean")) +
  geom_vline(aes(xintercept = median(rt_HYP),
                 color = "median")) +
  scale_color_manual(
    name = "Statistic",
    values = c("mean" = "blue", "median" = "red")
  ) +
  labs(x = "Syntactic complexity") +
  theme_minimal()

# boxplot
rt_HYP_box <- ggplot(dat, aes(y = rt_HYP)) +
  geom_boxplot() +
  labs(y = "Syntactic complexity") + 
  theme_minimal()

ppi = 200
png("plots/rt_HYP_uni_des.png", 
    width = 8*ppi, height = 4*ppi, res = ppi)
grid.arrange(rt_HYP_hist, rt_HYP_box, nrow = 1,
             top = "Distribution of syntactic complexity (rate of 'dat')")
dev.off()
```

```{r}
# summary stats
dat %>%
  summarise(
    mean = mean(rt_HYP),
    median = median(rt_HYP),
    sd = sd(rt_HYP),
    IQR = IQR(rt_HYP),
    skew = skewness(rt_HYP),
    kurtosis = kurtosis(rt_HYP)
  )
```

**Brouwer score**

```{r}
# histogram
bs_hist <- ggplot(dat_pca, aes(x = brouwer_score)) +
  geom_histogram() +
  geom_vline(aes(xintercept = mean(brouwer_score),
                 color = "mean")) +
  geom_vline(aes(xintercept = median(brouwer_score),
                 color = "median")) +
  scale_color_manual(
    name = "Statistic",
    values = c("mean" = "blue", "median" = "red")
  ) +
  labs(x = "Brouwer score") +
  theme_minimal()

# boxplot
bs_box <- ggplot(dat_pca, aes(y = brouwer_score)) +
  geom_boxplot() +
  labs(y = "Brouwer score") + 
  theme_minimal()

ppi = 200
png("plots/bs_uni_des.png", 
    width = 8*ppi, height = 4*ppi, res = ppi)
grid.arrange(bs_hist, bs_box, nrow = 1,
             top = "Distribution of Brouwer score")
dev.off()
```

```{r}
# summary stats
dat_pca %>%
  summarise(
    mean = mean(brouwer_score),
    median = median(brouwer_score),
    sd = sd(brouwer_score),
    IQR = IQR(brouwer_score),
    skew = skewness(brouwer_score),
    kurtosis = kurtosis(brouwer_score)
  )
```

**average number of connectives per sentence**

```{r}
# histogram
avg_conn_sent_hist <- ggplot(dat_pca, aes(x = avg_conn_sent)) +
  geom_histogram() +
  geom_vline(aes(xintercept = mean(avg_conn_sent),
                 color = "mean")) +
  geom_vline(aes(xintercept = median(avg_conn_sent),
                 color = "median")) +
  scale_color_manual(
    name = "Statistic",
    values = c("mean" = "blue", "median" = "red")
  ) +
  labs(x = "Average number of connectives per sentence") +
  theme_minimal()

# boxplot
avg_conn_sent_box <- ggplot(dat_pca, aes(y = avg_conn_sent)) +
  geom_boxplot() +
  labs(y = "Average number of connectives per sentence") + 
  theme_minimal()

ppi = 200
png("plots/avg_conn_sent_uni_des.png", 
    width = 8*ppi, height = 4*ppi, res = ppi)
grid.arrange(avg_conn_sent_hist, avg_conn_sent_box, nrow = 1,
             top = "Distribution of average number of connectives per sentence")
dev.off()
```

```{r}
# summary stats
dat %>%
  summarise(
    mean = mean(ratio_long_words),
    median = median(ratio_long_words),
    sd = sd(ratio_long_words),
    IQR = IQR(ratio_long_words),
    skew = skewness(ratio_long_words),
    kurtosis = kurtosis(ratio_long_words)
  )
```

**type-token ratio**

```{r}
# histogram
ttr_hist <- ggplot(dat_pca, aes(x = type_token_ratio)) +
  geom_histogram() +
  geom_vline(aes(xintercept = mean(type_token_ratio),
                 color = "mean")) +
  geom_vline(aes(xintercept = median(type_token_ratio),
                 color = "median")) +
  scale_color_manual(
    name = "Statistic",
    values = c("mean" = "blue", "median" = "red")
  ) +
  labs(x = "Type-token ratio") +
  theme_minimal()

# boxplot
ttr_box <- ggplot(dat_pca, aes(y = type_token_ratio)) +
  geom_boxplot() +
  labs(y = "Type-token ratio") + 
  theme_minimal()

ppi = 200
png("plots/ttr_uni_des.png", 
    width = 8*ppi, height = 4*ppi, res = ppi)
grid.arrange(ttr_hist, ttr_box, nrow = 1,
             top = "Distribution of type-token ratio")
dev.off()
```

```{r}
# summary stats
dat %>%
  summarise(
    mean = mean(type_token_ratio),
    median = median(type_token_ratio),
    sd = sd(type_token_ratio),
    IQR = IQR(type_token_ratio),
    skew = skewness(type_token_ratio),
    kurtosis = kurtosis(type_token_ratio)
  )
```

## 2.3. Bivariate Descriptives

**all numerical variables**

```{r}
# scatterplot matrix
ppi = 200
png("plots/pairsplot.png", 
    width = 14*ppi, height = 12*ppi, res = ppi)
ggpairs(dat, aes(color = lang4, alpha = 0.2),
        columns = c("nr_Ngr", "rt_EUS", "rt_LEX", "rt_HYP",
                    "brouwer_score", "avg_conn_sent",
                    "type_token_ratio")
        ) +
  theme_minimal()
dev.off()
```

**number of N-grams across varieties**

```{r}
# histograms
nr_Ngr_hist2 <- ggplot(dat, aes(x = nr_Ngr, fill = lang4)) +
  geom_histogram(alpha = 0.6) +
  facet_wrap(facet = ~ lang4) +
  labs(x = "Number of N-grams") +
  theme_minimal() +
  theme(legend.position = "none")

# boxplots
nr_Ngr_box2 <- ggplot(dat, aes(x = lang4, y = nr_Ngr, col = lang4)) +
  geom_boxplot() +
  labs(y = "Number of N-grams",
       x = "language variety") +
  theme_minimal() +
  theme(legend.position = "none")

ppi = 200
png("plots/nr_Ngr_bi_des.png", 
    width = 8*ppi, height = 7*ppi, res = ppi)
grid.arrange(nr_Ngr_hist2, nr_Ngr_box2, nrow = 2,
             top = "Distribution of number of N-grams per variety")
dev.off()
```

```{r}
# summary stats
dat %>% 
  group_by(lang4) %>%
  summarise(
    mean = mean(rt_FRM),
    median = median(rt_FRM),
    sd = sd(rt_FRM),
    IQR = IQR(rt_FRM),
    skew = skewness(rt_FRM),
    kurtosis = kurtosis(rt_FRM)
  )
```

**rate of *uh(m)* across varieties**

```{r}
# histograms
rt_EUS_hist2 <- ggplot(dat, aes(x = rt_EUS, fill = lang4)) +
  geom_histogram(alpha = 0.6) +
  facet_wrap(facet = ~ lang4) +
  labs(x = "Rate of 'uh(m)'") +
  theme_minimal() +
  theme(legend.position = "none")

# boxplots
rt_EUS_box2 <- ggplot(dat, aes(x = lang4, y = rt_EUS, col = lang4)) +
  geom_boxplot() +
  labs(y = "Rate of 'uh(m)'",
       x = "Language variety") +
  theme_minimal() +
  theme(legend.position = "none")

ppi = 200
png("plots/rt_EUS_bi_des.png", 
    width = 8*ppi, height = 7*ppi, res = ppi)
grid.arrange(rt_EUS_hist2, rt_EUS_box2, nrow = 2,
             top = "Distribution of disfluency (rate of 'uhm') per variety")
dev.off()
```

```{r}
# summary stats
dat %>%
  group_by(lang4) %>%
  summarise(
    mean = mean(rt_EUS),
    median = median(rt_EUS),
    sd = sd(rt_EUS),
    IQR = IQR(rt_EUS),
    skew = skewness(rt_EUS),
    kurtosis = kurtosis(rt_EUS)
  )
```

**lexical density across varieties**

```{r}
# histograms
rt_LEX_hist2 <- ggplot(dat, aes(x = rt_LEX, fill = lang4)) +
  geom_histogram(alpha = 0.6) +
  facet_wrap(facet = ~ lang4) +
  labs(x = "Lexical density") +
  theme_minimal() +
  theme(legend.position = "none")

# boxplots
rt_LEX_box2 <- ggplot(dat, aes(x = lang4, y = rt_LEX, col = lang4)) +
  geom_boxplot() +
  labs(y = "Lexical density",
       x = "Language variety") +
  theme_minimal() +
  theme(legend.position = "none")

ppi = 200
png("plots/rt_LEX_bi_des.png", 
    width = 8*ppi, height = 7*ppi, res = ppi)
grid.arrange(rt_LEX_hist2, rt_LEX_box2, nrow = 2,
             top = "Distribution of lexical density per variety")
dev.off()
```

```{r}
# summary stats
dat %>%
  group_by(lang4) %>%
  summarise(
    mean = mean(rt_LEX),
    median = median(rt_LEX),
    sd = sd(rt_LEX),
    IQR = IQR(rt_LEX),
    skew = skewness(rt_LEX),
    kurtosis = kurtosis(rt_LEX)
  )
```

**syntactic complexity across varieties**

```{r}
# histograms
rt_HYP_hist2 <- ggplot(dat, aes(x = rt_HYP, fill = lang4)) +
  geom_histogram(alpha = 0.6) +
  facet_wrap(facet = ~ lang4) +
  labs(x = "Syntactic complexity") +
  theme_minimal() +
  theme(legend.position = "none")

# boxplots
rt_HYP_box2 <- ggplot(dat, aes(x = lang4, y = rt_HYP, col = lang4)) +
  geom_boxplot() +
  labs(y = "Syntactic complexity",
       x = "Language variety") +
  theme_minimal() +
  theme(legend.position = "none")

ppi = 200
png("plots/rt_HYP_bi_des.png", 
    width = 8*ppi, height = 7*ppi, res = ppi)
grid.arrange(rt_HYP_hist2, rt_HYP_box2, nrow = 2,
             top = "Distribution of syntactic complexity (rate of 'dat') per variety")
dev.off()
```

```{r}
# summary stats
dat %>%
  group_by(lang4) %>%
  summarise(
    mean = mean(rt_HYP),
    median = median(rt_HYP),
    sd = sd(rt_HYP),
    IQR = IQR(rt_HYP),
    skew = skewness(rt_HYP),
    kurtosis = kurtosis(rt_HYP)
  )
```

**Brouwer score across varieties**

```{r}
# histograms
bs_hist2 <- ggplot(dat_pca, aes(x = brouwer_score, fill = lang4)) +
  geom_histogram(alpha = 0.6) +
  facet_wrap(facet = ~ lang4) +
  labs(x = "Brouwer score") +
  theme_minimal() +
  theme(legend.position = "none")

# boxplots
bs_box2 <- ggplot(dat_pca, aes(x = lang4, y = brouwer_score, col = lang4)) +
  geom_boxplot() +
  labs(y = "Brouwer score",
       x = "Language variety") +
  theme_minimal() +
  theme(legend.position = "none")

ppi = 200
png("plots/bs_bi_des.png", 
    width = 8*ppi, height = 7*ppi, res = ppi)
grid.arrange(bs_hist2, bs_box2, nrow = 2,
             top = "Distribution of Brouwer score per variety")
dev.off()
```

```{r}
# summary stats
dat_pca %>% 
  group_by(lang4) %>%
  summarise(
    mean = mean(brouwer_score),
    median = median(brouwer_score),
    sd = sd(brouwer_score),
    IQR = IQR(brouwer_score),
    skew = skewness(brouwer_score),
    kurtosis = kurtosis(brouwer_score)
  )
```

**average number of connectives per sentence across varieties**

```{r}
# histograms
avg_conn_sent_hist2 <- ggplot(dat_pca, aes(x = avg_conn_sent, fill = lang4)) +
  geom_histogram(alpha = 0.6) +
  facet_wrap(facet = ~ lang4) +
  labs(x = "Average number of connectives per sentence") +
  theme_minimal() +
  theme(legend.position = "none")

# boxplots
avg_conn_sent_box2 <- ggplot(dat_pca, aes(x = lang4, y = avg_conn_sent, 
                                          col = lang4)) +
  geom_boxplot() +
  labs(y = "Average number of connectives per sentence",
       x = "Language variety") +
  theme_minimal() +
  theme(legend.position = "none")

ppi = 200
png("plots/avg_conn_sent_bi_des.png", 
    width = 8*ppi, height = 7*ppi, res = ppi)
grid.arrange(avg_conn_sent_hist2, avg_conn_sent_box2, nrow = 2,
             top = "Distribution of average number of connectives per sentence per variety")
dev.off()
```

```{r}
# summary stats
dat_pca %>% 
  group_by(lang4) %>%
  summarise(
    mean = mean(avg_conn_sent),
    median = median(avg_conn_sent),
    sd = sd(avg_conn_sent),
    IQR = IQR(avg_conn_sent),
    skew = skewness(avg_conn_sent),
    kurtosis = kurtosis(avg_conn_sent)
  )
```

**type-token ratio across varieties**

```{r}
# histograms
ttr_hist2 <- ggplot(dat_pca, aes(x = type_token_ratio, fill = lang4)) +
  geom_histogram(alpha = 0.6) +
  facet_wrap(facet = ~ lang4) +
  labs(x = "Type-token ratio") +
  theme_minimal() +
  theme(legend.position = "none")

# boxplots
ttr_box2 <- ggplot(dat_pca, aes(x = lang4, y = type_token_ratio, col = lang4)) +
  geom_boxplot() +
  labs(y = "Type-token ratio",
       x = "Language variety") +
  theme_minimal() +
  theme(legend.position = "none")

ppi = 200
png("plots/ttr_bi_des.png", 
    width = 8*ppi, height = 7*ppi, res = ppi)
grid.arrange(ttr_hist2, ttr_box2, nrow = 2,
             top = "Distribution of type-token ratio per variety")
dev.off()
```

```{r}
# summary stats
dat_pca %>% 
  group_by(lang4) %>%
  summarise(
    mean = mean(type_token_ratio),
    median = median(type_token_ratio),
    sd = sd(type_token_ratio),
    IQR = IQR(type_token_ratio),
    skew = skewness(type_token_ratio),
    kurtosis = kurtosis(type_token_ratio)
  )
```

**table with all descriptives**

```{r}
# set variables
utterance_vars <- c("nr_Ngr", "rt_EUS", "rt_LEX", "rt_HYP")
text_vars <- c("brouwer_score", "avg_conn_sent", "type_token_ratio")

# function for descriptives
get_descriptives <- function(data, vars) {
  data %>%
    pivot_longer(
      cols = all_of(vars),
      names_to = "Variable",
      values_to = "Value"
    ) %>%
    group_by(Variable, lang4) %>%
    summarise(
      Mean     = round(mean(Value, na.rm = TRUE), 2),
      Median   = round(median(Value, na.rm = TRUE), 2),
      SD       = round(sd(Value, na.rm = TRUE), 2),
      IQR      = round(IQR(Value, na.rm = TRUE), 2),
      Skewness = round(skewness(Value, na.rm = TRUE), 2),
      Kurtosis = round(kurtosis(Value, na.rm = TRUE), 2),
      .groups  = "drop"
    )
}

# utterance-level descriptives
tab_utterance <- get_descriptives(dat, utterance_vars)

# text-level descriptives
tab_text <- get_descriptives(dat_pca, text_vars)

# combine
tab_descriptive <- bind_rows(
  tab_utterance,
  tab_text
)

# export
tab_descriptive %>%
  flextable() %>%
  autofit() %>%
  save_as_docx(path = "tables/descriptives.docx")
```

# 3. Modeling

## **3.1.2. Initial Model**

**Poisson model**

```{r}
# create offset and total number of N-grams per utterance
dat <- within(dat, off <- log(2 * nr_TOT - 5))

# model
m1 <- bam(nr_Ngr ~ lang4 + 
            s(rt_EUS, by = lang4) + 
            s(rt_LEX, by = lang4) +
            s(rt_HYP, by = lang4) + 
            s(brouwer_score, by = lang4) + 
            s(avg_conn_sent, by = lang4) + 
            s(type_token_ratio, by = lang4) +
            offset(off) + 
            s(fileID, bs = "re"),
          data = dat, family = poisson, discrete = TRUE, method = "fREML")

summary(m1)
```

```{r}
# initial plots
plot.gamViz(getViz(m1))
```

```{r}
# overdispersion check
pearson_chisq <- sum(resid(m1, type = "pearson")^2)
rdf <- df.residual(m1)
pearson_chisq / rdf
```

```{r}
# zero-inflation check
check_zeroinflation(m1)
```

```{r, message = FALSE}
# basis dimension check
k.check(m1)
```

```{r}
# default diagnostic plots
sink(tempfile())
check.gamViz(getViz(m1))
sink()
```

**negative binomial model**

```{r}
# refit m1 as negative binomial
m1_nb <- bam(nr_Ngr ~ lang4 + 
               s(rt_EUS, by = lang4) + 
               s(rt_LEX, by = lang4) +
               s(rt_HYP, by = lang4) + 
               s(brouwer_score, by = lang4) + 
               s(avg_conn_sent, by = lang4) + 
               s(type_token_ratio, by = lang4) +
               offset(off) + 
               s(fileID, bs = "re"),
             data = dat, family = nb, discrete = TRUE, method = "fREML")

AIC(m1, m1_nb) # m1_nb has much lower AIC
```

```{r}
# initial plots
plot.gamViz(getViz(m1_nb))
```

```{r}
# dispersion parameter: smaller = more overdispersion
m1_nb$family$getTheta(trans = TRUE)
```

```{r}
# zero-inflation check
check_zeroinflation(m1_nb)
```

```{r}
# basis dimension check
k.check(m1_nb)
```

```{r}
# default diagnostic plots
sink(tempfile())
check.gamViz(getViz(m1_nb))
sink()
```

```{r}
# concurivty
concurvity(m1_nb, full = TRUE) # severe concurvity
```

```{r}
# explore concurvity further

# pairwise concurvity measures
conc <- concurvity(m1_nb, full = FALSE)$estimate
round(conc, 2)

# pairwise correlations
vars <- c("rt_EUS", "rt_LEX", "rt_HYP", "brouwer_score",
          "avg_conn_sent", "type_token_ratio")
round(cor(dat[, vars]), 2)

# model comparisons
conc_test1 <- bam(nr_Ngr ~ s(brouwer_score) + offset(off) + 
                    s(fileID, bs = "re"),
                  data = dat, family = nb, discrete = TRUE, method = "fREML")
concurvity(conc_test1) # still high concurvity

conc_test2 <- bam(nr_Ngr ~ lang4 + 
                    s(rt_EUS, by = lang4) + 
                    s(rt_LEX, by = lang4) +
                    s(rt_HYP, by = lang4) + 
                    s(brouwer_score, by = lang4) + 
                    s(avg_conn_sent, by = lang4) + 
                    s(type_token_ratio, by = lang4) + 
                    offset(off),
                  data = dat, family = nb, discrete = TRUE, method = "fREML")

concurvity(conc_test2) # concurvity disappears when we remove random effect
summary(conc_test2)
plot.gamViz(getViz(conc_test2))
```

## 3.2. Model building

**exploring interactions**

```{r}
# stepwise selection of interactions: first step

# interaction between rt_EUS and rt_LEX
m2 <- bam(nr_Ngr ~ lang4 + 
            s(rt_EUS, by = lang4) + 
            s(rt_LEX, by = lang4) +
            ti(rt_EUS, rt_LEX, by = lang4) + 
            s(rt_HYP, by = lang4) +
            s(brouwer_score, by = lang4) + 
            s(avg_conn_sent, by = lang4) + 
            s(type_token_ratio, by = lang4) + 
            offset(off) +
            s(fileID, bs = "re"),
          data = dat, family = nb, discrete = TRUE, method = "fREML")

# interaction between rt_EUS and rt_HYP
m3 <- bam(nr_Ngr ~ lang4 + 
            s(rt_EUS, by = lang4) + 
            s(rt_LEX, by = lang4) +
            s(rt_HYP, by = lang4) + 
            ti(rt_EUS, rt_HYP, by = lang4) +
            s(brouwer_score, by = lang4) + 
            s(avg_conn_sent, by = lang4) + 
            s(type_token_ratio, by = lang4) + 
            offset(off) +
            s(fileID, bs = "re"),
          data = dat, family = nb, discrete = TRUE, method = "fREML")

# interaction between rt_EUS and brouwer_score
m4 <- bam(nr_Ngr ~ lang4 + 
            s(rt_EUS, by = lang4) + 
            s(rt_LEX, by = lang4) +
            s(rt_HYP, by = lang4) + 
            s(brouwer_score, by = lang4) + 
            ti(rt_EUS, brouwer_score, by = lang4) + 
            s(avg_conn_sent, by = lang4) +  
            s(type_token_ratio, by = lang4) + 
            offset(off) + 
            s(fileID, bs = "re"),
          data = dat, family = nb, discrete = TRUE, method = "fREML")

# interaction between rt_EUS and avg_conn_sent
m5 <- bam(nr_Ngr ~ lang4 + 
            s(rt_EUS, by = lang4) + 
            s(rt_LEX, by = lang4) +
            s(rt_HYP, by = lang4) + 
            s(brouwer_score, by = lang4) + 
            s(avg_conn_sent, by = lang4) + 
            ti(rt_EUS, avg_conn_sent, by = lang4) + 
            s(type_token_ratio, by = lang4) + 
            offset(off) + 
            s(fileID, bs = "re"),
          data = dat, family = nb, discrete = TRUE, method = "fREML")

# interaction between rt_EUS and type_token_ratio
m6 <- bam(nr_Ngr ~ lang4 + 
            s(rt_EUS, by = lang4) + 
            s(rt_LEX, by = lang4) +
            s(rt_HYP, by = lang4) + 
            s(brouwer_score, by = lang4) + 
            s(avg_conn_sent, by = lang4) + 
            s(type_token_ratio, by = lang4) + 
            ti(rt_EUS, type_token_ratio, by = lang4) + 
            offset(off) + 
            s(fileID, bs = "re"),
          data = dat, family = nb, discrete = TRUE, method = "fREML")

# comparisons
AIC(m1_nb, m2, m3, m4, m5, m6) # m3 preferred
```

```{r}
# stepwise selection of interactions: second step

# interaction between rt_EUS and rt_LEX
m7 <- bam(nr_Ngr ~ lang4 + 
            s(rt_EUS, by = lang4) + 
            s(rt_LEX, by = lang4) + 
            ti(rt_EUS, rt_LEX, by = lang4) + 
            s(rt_HYP, by = lang4) + 
            ti(rt_EUS, rt_HYP, by = lang4) + 
            s(brouwer_score, by = lang4) + 
            s(avg_conn_sent, by = lang4) + 
            s(type_token_ratio, by = lang4) + 
            offset(off) + 
            s(fileID, bs = "re"),
          data = dat, family = nb, discrete = TRUE, method = "fREML")

# interaction between rt_EUS and brouwer_score
m8 <- bam(nr_Ngr ~ lang4 + 
            s(rt_EUS, by = lang4) + 
            s(rt_LEX, by = lang4) +
            s(rt_HYP, by = lang4) + 
            ti(rt_EUS, rt_HYP, by = lang4) +
            s(brouwer_score, by = lang4) + 
            ti(rt_EUS, brouwer_score, by = lang4) + 
            s(avg_conn_sent, by = lang4) + 
            s(type_token_ratio, by = lang4) + 
            offset(off) + 
            s(fileID, bs = "re"),
          data = dat, family = nb, discrete = TRUE, method = "fREML")

# interaction between rt_EUS and avg_conn_sent
m9 <- bam(nr_Ngr ~ lang4 + 
            s(rt_EUS, by = lang4) + 
            s(rt_LEX, by = lang4) +
            s(rt_HYP, by = lang4) + 
            ti(rt_EUS, rt_HYP, by = lang4) +
            s(brouwer_score, by = lang4) + 
            s(avg_conn_sent, by = lang4) + 
            ti(rt_EUS, avg_conn_sent, by = lang4) + 
            s(type_token_ratio, by = lang4) +  
            offset(off) + 
            s(fileID, bs = "re"),
          data = dat, family = nb, discrete = TRUE, method = "fREML")

# interaction between rt_EUS and type_token_ratio
m10 <- bam(nr_Ngr ~ lang4 + 
             s(rt_EUS, by = lang4) + 
             s(rt_LEX, by = lang4) +
             s(rt_HYP, by = lang4) + 
             ti(rt_EUS, rt_HYP, by = lang4) +
             s(brouwer_score, by = lang4) + 
             s(avg_conn_sent, by = lang4) + 
             s(type_token_ratio, by = lang4) + 
             ti(rt_EUS, type_token_ratio, by = lang4) + 
             offset(off) + 
             s(fileID, bs = "re"),
          data = dat, family = nb, discrete = TRUE, method = "fREML")

# comparisons
AIC(m3, m7, m8, m9, m10) # m10 preferred
```

```{r}
# stepwise selection of interactions: third step

# interaction between rt_EUS and rt_LEX
m11 <- bam(nr_Ngr ~ lang4 + 
             s(rt_EUS, by = lang4) + 
             s(rt_LEX, by = lang4) +
             ti(rt_EUS, rt_LEX, by = lang4) + 
             s(rt_HYP, by = lang4) + 
             ti(rt_EUS, rt_HYP, by = lang4) + 
             s(brouwer_score, by = lang4) + 
             s(avg_conn_sent, by = lang4) + 
             s(type_token_ratio, by = lang4) + 
             ti(rt_EUS, type_token_ratio, by = lang4) + 
             offset(off) + 
             s(fileID, bs = "re"),
           data = dat, family = nb, discrete = TRUE, method = "fREML")

# interaction between rt_EUS and brouwer_score
m12 <- bam(nr_Ngr ~ lang4 + 
             s(rt_EUS, by = lang4) + 
             s(rt_LEX, by = lang4) +
             s(rt_HYP, by = lang4) + 
             ti(rt_EUS, rt_HYP, by = lang4) +
             s(brouwer_score, by = lang4) + 
             ti(rt_EUS, brouwer_score, by = lang4) + 
             s(avg_conn_sent, by = lang4) + 
             s(type_token_ratio, by = lang4) +
             ti(rt_EUS, type_token_ratio, by = lang4) + 
             offset(off) + 
             s(fileID, bs = "re"),
           data = dat, family = nb, discrete = TRUE, method = "fREML")

# interaction between rt_EUS and avg_conn_sent
m13 <- bam(nr_Ngr ~ lang4 + 
             s(rt_EUS, by = lang4) + 
             s(rt_LEX, by = lang4) +
             s(rt_HYP, by = lang4) + 
             ti(rt_EUS, rt_HYP, by = lang4) +
             s(brouwer_score, by = lang4) + 
             s(avg_conn_sent, by = lang4) + 
             ti(rt_EUS, avg_conn_sent, by = lang4) + 
             s(type_token_ratio, by = lang4) + 
             ti(rt_EUS, type_token_ratio, by = lang4) + 
             offset(off) + 
             s(fileID, bs = "re"),
           data = dat, family = nb, discrete = TRUE, method = "fREML")

# comparisons
AIC(m10, m11, m12, m13) # m13 preferred
```

```{r}
# stepwise selection of interactions: fourth step

# interaction between rt_EUS and rt_LEX
m14 <- bam(nr_Ngr ~ lang4 + 
             s(rt_EUS, by = lang4) + 
             s(rt_LEX, by = lang4) +
             ti(rt_EUS, rt_LEX, by = lang4) + 
             s(rt_HYP, by = lang4) + 
             ti(rt_EUS, rt_HYP, by = lang4) + 
             s(brouwer_score, by = lang4) + 
             s(avg_conn_sent, by = lang4) + 
             ti(rt_EUS, avg_conn_sent, by = lang4) + 
             s(type_token_ratio, by = lang4) + 
             ti(rt_EUS, type_token_ratio, by = lang4) + 
             offset(off) + 
             s(fileID, bs = "re"),
           data = dat, family = nb, discrete = TRUE, method = "fREML")

# interaction between rt_EUS and brouwer_score
m15 <- bam(nr_Ngr ~ lang4 + 
             s(rt_EUS, by = lang4) + 
             s(rt_LEX, by = lang4) +
             s(rt_HYP, by = lang4) + 
             ti(rt_EUS, rt_HYP, by = lang4) +
             s(brouwer_score, by = lang4) + 
             ti(rt_EUS, brouwer_score, by = lang4) +
             s(avg_conn_sent, by = lang4) + 
             ti(rt_EUS, avg_conn_sent, by = lang4) + 
             s(type_token_ratio, by = lang4) + 
             ti(rt_EUS, type_token_ratio, by = lang4) +
             offset(off) + 
             s(fileID, bs = "re"),
           data = dat, family = nb, discrete = TRUE, method = "fREML")

# comparisons
AIC(m13, m14, m15) # no further improvement > keep m13
```

**checking selected model**

```{r}
# intitial plots
plot.gamViz(getViz(m13))
```

```{r}
# EDF and basis dimension check
round(k.check(m13), 2)
```

```{r}
# zero-inflation check
check_zeroinflation(m10)
```

```{r}
# default diagnostic plots
ppi = 200
png("plots/m13_diagnostics.png", 
    width = 9*ppi, height = 7*ppi, res = ppi)
sink(tempfile())
check.gamViz(getViz(m10))
sink()
dev.off()
```

```{r}
# normal QQ-plot of random intercepts
ppi = 200
png("plots/m13_QQ_ranef.png", 
    width = 9*ppi, height = 7*ppi, res = ppi)
plot(sm(getViz(m10), 25))
dev.off()
```

```{r}
# diagnostic plots per variety

# extract residuals and fitted values
dat <- within(dat, resids <- resid(m13, type = "deviance"))
dat <- within(dat, fitted <- fitted(m13))
dat <- within(dat, lin_pred <- m13$linear.predictors)

# variety labels
variety_labs <- c(`NED_in` = "Interpreted Dutch",
                  `NED_nl` = "Netherlandic Dutch",
                  `NED_vl` = "Belgian Dutch")

# residuals vs. linear predictor per variety
ppi = 200
png("plots/m13_res_vs_linpred.png", 
    width = 9*ppi, height = 7*ppi, res = ppi)
ggplot(dat, aes(x = lin_pred, y = resids)) +
  geom_point(alpha = 0.5) +
  facet_wrap(facets = ~ lang4, nrow = 2,
             labeller = as_labeller(variety_labs)) +
  labs(x = "Residuals",
       y = "Linear predictor") +
  theme_minimal()
dev.off()

# histogram of residuals per variety
ppi = 200
png("plots/m13_hist_res.png", 
    width = 9*ppi, height = 7*ppi, res = ppi)
ggplot(dat, aes(x = resids)) +
  geom_histogram() +
  facet_wrap(facets = ~ lang4, nrow = 2,
             labeller = as_labeller(variety_labs)) +
  labs(x = "Residuals",
       y = "Count") +
  theme_minimal()
dev.off()

# response vs. fitted values per variety
cor_dat <- dat %>% group_by(lang4) %>% 
  summarise(r = cor(nr_Ngr, fitted))

ppi = 200
png("plots/m13_resp_vs.fit.png", 
    width = 9*ppi, height = 7*ppi, res = ppi)
ggplot(dat, aes(x = fitted, y = nr_Ngr)) +
  geom_point(alpha = 0.5) +
  facet_wrap(facets = ~ lang4, nrow = 2,
             labeller = as_labeller(variety_labs)) +
  labs(x = "Response",
       y = "Fitted values") +
  geom_text(
    data = cor_dat, aes(x = 8, y = 11,
                        label = paste0("r = ", round(r, 2))),
    hjust = 1.1, vjust = 1.5) +
  theme_minimal()
dev.off()
```

```{r}
# index plot of residuals to check outliers
ppi = 200
png("plots/m13_res_index.png", 
    width = 9*ppi, height = 5*ppi, res = ppi)
ggplot(dat, aes(x = 1:length(resids), y = resids, col = lang4)) +
  geom_point(alpha = 0.5) +
  labs(x = "Observation index",
       y = "Residuals",
       col = "Variety") +
  scale_color_discrete(labels = c("Interpreted Dutch",
                                  "Netherlandic Dutch",
                                  "Belgian Dutch")) +
  theme_minimal()
dev.off()
```

```{r}
# check observations with outlying residuals
dat[which(dat$resids > 2.5), ]

# refit model without these observations and compare
dat_red <- dat[-which(dat$resids > 2.5), ]

m13_red <- bam(nr_Ngr ~ lang4 + 
                 s(rt_EUS, by = lang4) + 
                 s(rt_LEX, by = lang4) +
                 s(rt_HYP, by = lang4) + 
                 ti(rt_EUS, rt_HYP, by = lang4) +
                 s(brouwer_score, by = lang4) + 
                 s(avg_conn_sent, by = lang4) + 
                 ti(rt_EUS, avg_conn_sent, by = lang4) + 
                 s(type_token_ratio, by = lang4) + 
                 ti(rt_EUS, type_token_ratio, by = lang4) + 
                 offset(off) + 
                 s(fileID, bs = "re"),
               data = dat_red, family = nb, discrete = TRUE, method = "fREML")

summary(m13_red) # no substantial changes
plot.gamViz(getViz(m13_red))
check.gamViz(getViz(m13_red))
cor(fitted(m13_red), dat_red$nr_Ngr)
```

```{r}
# residuals vs. predictor plots

# residuals vs. rt_EUS
ppi = 200
png("plots/m13_res_vs_rt_EUS.png", 
    width = 9*ppi, height = 7*ppi, res = ppi)
ggplot(dat, aes(x = rt_EUS, y = resids)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(x = "Disfluency (rate of 'uhm')",
       y = "Residuals") +
  theme_minimal()
dev.off()

# residuals vs. rt_LEX
ppi = 200
png("plots/m13_res_vs_rt_LEX.png", 
    width = 9*ppi, height = 7*ppi, res = ppi)
ggplot(dat, aes(x = rt_LEX, y = resids)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(x = "Lexical density",
       y = "Residuals") +
  theme_minimal()
dev.off()

# residuals vs. rt_HYP
ppi = 200
png("plots/m13_res_vs_rt_HYP.png", 
    width = 9*ppi, height = 7*ppi, res = ppi)
ggplot(dat, aes(x = rt_HYP, y = resids)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(x = "Syntactic complexity (rate of 'dat')",
       y = "Residuals") +
  theme_minimal()
dev.off()

# residuals vs. brouwer_score
ppi = 200
png("plots/m13_res_vs_bs.png", 
    width = 9*ppi, height = 7*ppi, res = ppi)
ggplot(dat, aes(x = brouwer_score, y = resids)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(x = "Brouwer score",
       y = "Residuals") +
  theme_minimal()
dev.off()

# residuals vs. avg_conn_sent
ppi = 200
png("plots/m13_res_avg_conn_sent.png", 
    width = 9*ppi, height = 7*ppi, res = ppi)
ggplot(dat, aes(x = avg_conn_sent, y = resids)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(x = "avg_conn_sent",
       y = "Residuals") +
  theme_minimal()
dev.off()

# residuals vs. type_token_ratio
ppi = 200
png("plots/m13_res_vs_ttr.png", 
    width = 9*ppi, height = 7*ppi, res = ppi)
ggplot(dat, aes(x = type_token_ratio, y = resids)) +
  geom_point() +
  geom_smooth(method = "loess") +
  labs(x = "Type-token ratio",
       y = "Residuals") +
  theme_minimal()
dev.off()
```

```{r}
# concurvity
concurvity(m13, full = TRUE)
```

```{r}
# explore concurvity further

# pairwise concurvity measures
conc2 <- concurvity(m13, full = FALSE)$estimate
round(conc2, 2)

# compare with model without random effect
conc_test3 <- bam(nr_Ngr ~ lang4 + 
                    s(rt_EUS, by = lang4) + 
                    s(rt_LEX, by = lang4) +
                    s(rt_HYP, by = lang4) + 
                    ti(rt_EUS, rt_HYP, by = lang4) +
                    s(brouwer_score, by = lang4) + 
                    s(avg_conn_sent, by = lang4) + 
                    ti(rt_EUS, avg_conn_sent, by = lang4) + 
                    s(type_token_ratio, by = lang4) + 
                    ti(rt_EUS, type_token_ratio, by = lang4) + 
                    offset(off),
                  data = dat, family = nb, discrete = TRUE, method = "fREML")

concurvity(conc_test3)
```

## 3.3. Final model interpretation

```{r}
# summary table
summary(m13)

# compare with deviance explained of null model
null_mod <- bam(nr_Ngr ~ offset(off) + s(fileID, bs = "re"),
                data = dat, family = nb, method = "fREML")
summary(null_mod)
```

```{r}
# export summary table
m13_tbl <- as_flextable(m13) %>%
  colformat_double(digits = 3) %>%
  set_table_properties(layout = "fixed") %>%
  fit_to_width(max_width = 6.5)

save_as_docx(
  "GAM Model Summary" = m13_tbl, 
  path = "tables/m13_summary.docx"
)
```

```{r}
# effect plot lang4
ppi = 200
png("plots/effect_lang4.png", 
    width = 7*ppi, height = 5*ppi, res = ppi)
plot(predict_response(m13, terms = "lang4")) +
  geom_line() +
  labs(x = "Variety",
       y = "Predicted rate of N-grams",
       title = "")
dev.off()
```

```{r}
# effect plot rt_EUS
ppi = 200
png("plots/effect_rt_EUS.png", 
    width = 9*ppi, height = 5*ppi, res = ppi)
visreg(m13, "rt_EUS", by = "lang4",
       scale = "response", rug = TRUE) +
  coord_cartesian(ylim = c(0, 1)) +
  labs(x = "Disfluency (rate of 'uhm')",
       y = "Predicted rate of N-grams") +
  facet_wrap(facets = ~ lang4, nrow = 1,
             labeller = as_labeller(variety_labs)) +
  theme_minimal()
dev.off()
```

```{r}
# effect plot rt_LEX
ppi = 200
png("plots/effect_rt_LEX.png", 
    width = 9*ppi, height = 5*ppi, res = ppi)
visreg(m13, "rt_LEX", by = "lang4",
       scale = "response", rug = TRUE) +
  coord_cartesian(ylim = c(0, 8)) +
  labs(x = "Lexical density",
       y = "Predicted rate of N-grams") +
  facet_wrap(facets = ~ lang4, nrow = 1,
             labeller = as_labeller(variety_labs)) +
  theme_minimal()
dev.off()
```

```{r}
# effect plot rt_HYP
ppi = 200
png("plots/effect_rt_HYP.png", 
    width = 9*ppi, height = 5*ppi, res = ppi)
visreg(m13, "rt_HYP", by = "lang4",
       scale = "response", rug = TRUE) +
  coord_cartesian(ylim = c(0, 2.5)) +
  labs(x = "Syntactic complexity (rate of 'dat')",
       y = "Predicted rate of N-grams") +
  facet_wrap(facets = ~ lang4, nrow = 1,
             labeller = as_labeller(variety_labs)) +
  theme_minimal()
dev.off()
```

```{r}
# effect plot brouwer_score
ppi = 200
png("plots/effect_bs.png", 
    width = 9*ppi, height = 5*ppi, res = ppi)
visreg(m13, "brouwer_score", by = "lang4",
       scale = "response", rug = TRUE) +
  coord_cartesian(ylim = c(0, 2)) +
  labs(x = "Brouwer score",
       y = "Predicted rate of N-grams") +
  facet_wrap(facets = ~ lang4, nrow = 1,
             labeller = as_labeller(variety_labs)) +
  theme_minimal()
dev.off()
```

```{r}
# effect plot avg_conn_sent
ppi = 200
png("plots/effect_avg_conn_sent.png", 
    width = 9*ppi, height = 5*ppi, res = ppi)
visreg(m13, "avg_conn_sent", by = "lang4",
       scale = "response", rug = TRUE) +
  coord_cartesian(ylim = c(0, 1.5)) +
  labs(x = "Average number of connectives per sentence",
       y = "Predicted rate of N-grams") +
  facet_wrap(facets = ~ lang4, nrow = 1,
             labeller = as_labeller(variety_labs)) +
  theme_minimal()
dev.off()
```

```{r}
# effect plot type_token_ratio
ppi = 200
png("plots/effect_TTR.png", 
    width = 9*ppi, height = 5*ppi, res = ppi)
visreg(m13, "type_token_ratio", by = "lang4",
       scale = "response", rug = TRUE) +
  coord_cartesian(ylim = c(0, 2)) +
  labs(x = "Type-token ratio",
       y = "Predicted rate of N-grams") +
  facet_wrap(facets = ~ lang4, nrow = 1,
             labeller = as_labeller(variety_labs)) +
  theme_minimal()
dev.off()
```

```{r}
# effect plots rt_EUS x rt_HYP

# NED_in
rt_EUS_rt_HYP_ID <- visreg2d(m13, "rt_EUS", "rt_HYP", scale = "response",
                             cond = list(lang4 = "NED_in")) +
  labs(x = "Disfluency",
       y = "Syntactic complexity",
       title = "Interpreted Dutch")

# NED_nl
rt_EUS_rt_HYP_ND <- visreg2d(m13, "rt_EUS", "rt_HYP", scale = "response",
                             cond = list(lang4 = "NED_nl")) +
  labs(x = "Disfluency",
       y = "Syntactic complexity",
       title = "Netherlandic Dutch")

# NED_vl
rt_EUS_rt_HYP_BD <- visreg2d(m13, "rt_EUS", "rt_HYP", scale = "response",
                             cond = list(lang4 = "NED_vl")) +
  labs(x = "Disfluency",
       y = "Syntactic complexity",
       title = "Belgian Dutch")

# combine
ppi = 200
png("plots/effect_rt_EUS_rt_HYP.png", 
    width = 10*ppi, height = 6*ppi, res = ppi)
grid.arrange(rt_EUS_rt_HYP_ID, rt_EUS_rt_HYP_ND,
             rt_EUS_rt_HYP_BD, nrow = 2)
dev.off()
```

```{r}
# effect plots rt_EUS x type_token_ratio

# NED_in
rt_EUS_TTR_ID <- visreg2d(m13, "rt_EUS", "type_token_ratio", scale = "response",
                          cond = list(lang4 = "NED_in")) +
  labs(x = "Disfluency (rate of 'uhm')",
       y = "Type-token ratio",
       title = "Interpreted Dutch")

# NED_nl
rt_EUS_TTR_ND <- visreg2d(m13, "rt_EUS", "type_token_ratio", scale = "response",
                          cond = list(lang4 = "NED_nl")) +
  labs(x = "Disfluency (rate of 'uhm')",
       y = "Type-token ratio",
       title = "Netherlandic Dutch")

# NED_vl
rt_EUS_TTR_BD <- visreg2d(m13, "rt_EUS", "type_token_ratio", scale = "response",
                          cond = list(lang4 = "NED_vl")) +
  labs(x = "Disfluency (rate of 'uhm')",
       y = "Type-token ratio",
       title = "Belgian Dutch")

# combine
ppi = 200
png("plots/effect_rt_EUS_TTR.png", 
    width = 10*ppi, height = 6*ppi, res = ppi)
grid.arrange(rt_EUS_TTR_ID, rt_EUS_TTR_ND,
             rt_EUS_TTR_BD, nrow = 2)
dev.off()
```

```{r}
# effect plots rt_EUS x avg_conn_sent

# NED_in
rt_EUS_TTR_ID <- visreg2d(m13, "rt_EUS", "avg_conn_sent", scale = "response",
                          cond = list(lang4 = "NED_in")) +
  labs(x = "Disfluency (rate of 'uhm')",
       y = "Average connectives per sentence",
       title = "Interpreted Dutch")

# NED_nl
rt_EUS_TTR_ND <- visreg2d(m13, "rt_EUS", "avg_conn_sent", scale = "response",
                          cond = list(lang4 = "NED_nl")) +
  labs(x = "Disfluency (rate of 'uhm')",
       y = "Average connectives per sentence",
       title = "Netherlandic Dutch")

# NED_vl
rt_EUS_TTR_BD <- visreg2d(m13, "rt_EUS", "avg_conn_sent", scale = "response",
                          cond = list(lang4 = "NED_vl")) +
  labs(x = "Disfluency (rate of 'uhm')",
       y = "Average connectives per sentence",
       title = "Belgian Dutch")

# combine
ppi = 200
png("plots/effect_rt_EUS_avg_conn_sent.png", 
    width = 10*ppi, height = 6*ppi, res = ppi)
grid.arrange(rt_EUS_TTR_ID, rt_EUS_TTR_ND,
             rt_EUS_TTR_BD, nrow = 2)
dev.off()
```
